A cautionary note on the use of Ornstein Uhlenbeck models in macroevolutionary studies

### 2015-05-11 Manuscript/ohyou.tex
Fixed referencing errors

    We find that Type I error rates are unacceptably high when tree size is small (Tables \ref{tables:bd_noerror} and \ref{tables:varrate_noerror}), i.e., the OU model is often favoured, even though the Brownian model was used to generate the data. 
    For some tree shapes, particularly where speciation rates accelerate towards the present, Type I error remains \textgreater 0.05 even for trees with 1000 tips (Table \ref{tables:varrate_noerror}). 
    Indeed, for such trees the confidence limits on parameter estimates are broad (\ref{figure:ouprofiles})
    This shows that, in general, analyses based on small datasets are prone to biases that decrease only slowly as the size of the dataset increases. 
    Unfortunately, OU models are often fitted to phylogenies with fewer than 100 taxa (mean = 166.97 $\pm$ 43.86, median = 58; Figure \ref{figure:ntaxa}; see Supporting Information).

@@

  \subsection{Recommendations for interpreting $\alpha$}    
    \begin{enumerate}[(i)]
    \item \textbf{Consider plausible alternative hypotheses.}
    Because the OU model was proposed with evolutionary process in mind, alternative and arguably more parsimonious explanations for favouring OU models and interpreting non-zero $\alpha$ are often overlooked. 
    The effects of measurement error in particular suggest that $\alpha$ must always be interpreted with caution. 
    Many issues of misinterpretation can be solved by carefully inspecting the $\alpha$ parameter when an OU model is favoured. 
    Often, when Likelihood ratio tests suggest the OU model should be favoured over the Brownian model, estimates of $\alpha$ are actually very close to zero, i.e. very close to Brownian.
    In these circumstances, it is likely that measurement error, or similar, is being mopped up by the extra parameters in the OU model. 
    Thus this does not reflect any kind of OU process underlying the data. 
    The similarity between Brownian and OU models with small $\alpha$ is demonstrated in (\ref{figure:ouhistory}). 
    At the other extreme, as values of $\alpha$ become larger, the effects of changing $\alpha$ on model predictions are increasingly small and large values of $\alpha$ are indistinguishable from white-noise.

>>>>

    We find that Type I error rates are unacceptably high when tree size is small (Tables \ref{tables:bd_noerror} and \ref{tables:varrate_noerror}), i.e., the OU model is often favoured, even though the Brownian model was used to generate the data. 
    For some tree shapes, particularly where speciation rates accelerate towards the present, Type I error remains \textgreater 0.05 even for trees with 1000 tips (Table \ref{tables:varrate_noerror}). 
    Indeed, for such trees the confidence limits on parameter estimates are broad (Figure \ref{figure:ouprofiles})
    This shows that, in general, analyses based on small datasets are prone to biases that decrease only slowly as the size of the dataset increases. 
    Unfortunately, OU models are often fitted to phylogenies with fewer than 100 taxa (mean = 166.97 $\pm$ 43.86, median = 58; Figure \ref{figure:ntaxa}; see Supporting Information).

@@

  \subsection{Recommendations for interpreting $\alpha$}    
    \begin{enumerate}[(i)]
    \item \textbf{Consider plausible alternative hypotheses.}
    Because the OU model was proposed with evolutionary process in mind, alternative and arguably more parsimonious explanations for favouring OU models and interpreting non-zero $\alpha$ are often overlooked. 
    The effects of measurement error in particular suggest that $\alpha$ must always be interpreted with caution. 
    Many issues of misinterpretation can be solved by carefully inspecting the $\alpha$ parameter when an OU model is favoured. 
    Often, when Likelihood ratio tests suggest the OU model should be favoured over the Brownian model, estimates of $\alpha$ are actually very close to zero, i.e. very close to Brownian.
    In these circumstances, it is likely that measurement error, or similar, is being mopped up by the extra parameters in the OU model. 
    Thus this does not reflect any kind of OU process underlying the data. 
    The similarity between Brownian and OU models with small $\alpha$ is demonstrated in (Figure \ref{figure:ouhistory}). 
    At the other extreme, as values of $\alpha$ become larger, the effects of changing $\alpha$ on model predictions are increasingly small and large values of $\alpha$ are indistinguishable from white-noise.
